Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-08-27 draft.tex
Some followup on Sec. IV

It remains to invert the relation \refeq{linear-assumption}, so that one can determine the unfolded
angular observables $\lbrace \wwhat{S_k}\rbrace$ as
\begin{equation}
    \wwhat{S_i} = \danny{N_\eps} \left(M^{-1}\right)_{ij} \est{S_j}\,.
\end{equation}
\danny{Here $N_\eps$ is a normalization factor based on the detector acceptance function $\eps$. It emerges
    since the product $P(x) \eps(x)$ is not a PDF; it is not normalized appropriately\footnote{We thank
    X.X\danny{Marcin: Please insert whathisname} for pointing this out.}. We can determine $N_\eps$ by
    imposing
    \begin{gather}
        1 \equiv \int \rmdx{x} P(x | \lbrace \wwhat{S}_k\rbrace) = N_\eps \sum_{k,j} \left(M^{-1}\right)_{kj} \est{S_j}\int \rmdx{x} f_k(x)\,,\\
        \Rightarrow N_\eps \equiv \frac{1}{\sum_{k,j} \left(M^{-1}\right)_{kj} \est{S_j} \int \rmdx{x} f_k(x)}\,.
    \end{gather}
}

@@

\fred{$N$ used for the symbol $K$ of section I and II}
Our finding can be summarized as follows:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
        All pull distributions can successfully be fitted to a gaussian distribution. As an example, we
        show the pull distributions for the observables $S_5$ and $S_7$ for a SM-like $B\to K^*\ell^+\ell^-$
        decays in \reffig{pulls}.
    \item We study the absolute uncertainty $\sigma_k(K)$ with respect to the angular observable $S_k$
        as a function of the number of simulated events $K$. As expected for a multivariate gaussian
        distribution, we find that the absolute uncertainty is well fitted by
        \begin{equation}
            \sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$, regardless of the absolute size of $S_k$. The latter can best be shown
        for the uncertainties of two exampular observables. Taking again $S_5$ ($\simeq 27\%$) and $S_7$ ($\simeq 2\%$)
        for SM-like $B\to K^*\ell^+\ell^-$ decays, we show the absolute uncertainty in \reffig{errors}.
    \item For all toy studies, we produce plots of their skewness and their kurtosis as functions of
        $N$. We find in all cases that the mean skewness and kurtosis converge towards zero.
        \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of the number of simulated events $K$.}
\end{itemize}
We find that all of the above results show consistently that the joint distrbution of the angular observables
converges towards a multivariate Gaussian distribution. We propose to describe the resulting distribution as
a multivariate Gauss when skewness and kurtosis are small enough. The precise targets for these quantities
are subject to the details of the physical analysis, and therefore outside the scope of our work.

@@

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/error-Q2_5_6_S5.png}
            \includegraphics[width=0.45\textwidth]{figs/error-Q2_5_6_S7.png}
        \caption{Uncertainty of the angular observables $S_5$ and $S_7$, extracted from simulated events of the decay $B\to K^*\ell^+\ell^-$ and as a function of the number of simulated events $N$. The red curve represents a fit to the function $\sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}$.}
        \label{fig:errors}
\end{figure}

{\textbf{[dvd: can the following be removed?]}
We used the simulated events and fitted a likelihood function to them.
\begin{itemize}
    \item For a small number of simulated events, \danny{$N=?$}, we observe a significant bias in the pull distributions.
    \item For a large number of simulated events, \danny{$N=?$}, the uncertainty of the angular observables is
        smaller compared to the results of the method of moments.
\end{itemize}
\todo{compare robustness to likelihood fit when fit doesn't
converge. Why exactly doesn't it converge? What if physics model
wrong, say some angular terms are missing.}}

>>>>

It remains to invert the relation \refeq{linear-assumption}, so that one can determine the unfolded
angular observables $\lbrace \wwhat{S_k}\rbrace$ as
\begin{equation}
    \wwhat{S_i} = \danny{N_\eps} \left(M^{-1}\right)_{ij} \est{S_j}\,.
\end{equation}
\danny{Here $N_\eps$ is a normalization factor based on the detector acceptance function $\eps$. It emerges
    since the product $P(x) \eps(x)$ is not a PDF; it is not normalized appropriately\footnote{We thank
    X.X\danny{Marcin: Please insert whatshisname} for pointing this out.}. We can determine $N_\eps$ by
    imposing
    \begin{gather}
        1 \equiv \int \rmdx{x} P(x | \lbrace \wwhat{S}_k\rbrace) = N_\eps \sum_{k,j} \left(M^{-1}\right)_{kj} \est{S_j}\int \rmdx{x} f_k(x)\,,\\
        \Rightarrow N_\eps \equiv \frac{1}{\sum_{k,j} \left(M^{-1}\right)_{kj} \est{S_j} \int \rmdx{x} f_k(x)}\,.
    \end{gather}
}

@@

Our finding can be summarized as follows:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
        All pull distributions can successfully be fitted to a gaussian distribution. As an example, we
        show the pull distributions for the observables $S_5$ and $S_7$ for a SM-like $B\to K^*\ell^+\ell^-$
        decays in \reffig{pulls}.
    \item We study the absolute uncertainty $\sigma_k(K)$ with respect to the angular observable $S_k$
        as a function of the number of simulated events $K$. As expected for a multivariate gaussian
        distribution, we find that the absolute uncertainty is well fitted by
        \begin{equation}
            \label{eq:unc-on-mean}
            \sigma_k(K) = \frac{\sigma_k(1)}{\sqrt{K}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$, regardless of the absolute size of $S_k$. The latter can best be shown
        for the uncertainties of two exampular observables. Taking again $S_5$ ($\simeq 27\%$) and $S_7$ ($\simeq 2\%$)
        for SM-like $B\to K^*\ell^+\ell^-$ decays, we show the absolute uncertainty in \reffig{errors}.
    \item For all toy studies, we produce plots of their skewness and their kurtosis as functions of
        $N$. We find in all cases that the mean skewness and kurtosis converge towards zero.
        \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of the number of simulated events $K$.}
\end{itemize}
We find that all of the above results show consistently that the joint distrbution of the angular observables
converges towards a multivariate Gaussian distribution. We propose to describe the resulting distribution as
a multivariate Gauss when skewness and kurtosis are small enough. The precise targets for these quantities
are subject to the details of the physical analysis, and therefore outside the scope of our work.

@@

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/error-Q2_5_6_S5.png}
            \includegraphics[width=0.45\textwidth]{figs/error-Q2_5_6_S7.png}
            \caption{Uncertainty of the angular observables $S_5$ and $S_7$, extracted from simulated events of the decay $B\to K^*\ell^+\ell^-$ and as a function of the number of simulated events $K$. The red curve represents a fit to the function given in \refeq{unc-on-mean}.}
        \label{fig:errors}
\end{figure}

\hrulefill\\
{\textbf{\color{red}[dvd: can the following be removed?]}
We used the simulated events and fitted a likelihood function to them.
\begin{itemize}
    \item For a small number of simulated events, \danny{$K=?$}, we observe a significant bias in the pull distributions.
    \item For a large number of simulated events, \danny{$K=?$}, the uncertainty of the angular observables is
        smaller compared to the results of the method of moments.
\end{itemize}
todo: compare robustness to likelihood fit when fit doesn't
converge. Why exactly doesn't it converge? What if physics model
wrong, say some angular terms are missing.}
\hrulefill
